Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States

### 2024-11-01 paper/paper.qmd
fixed a little prose

While individual polls provide snapshots of public opinion, they are often subject to biases and methodological differences. In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by aggregating data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy. Our analysis focuses on estimating the levels of support for each candidate not only nationally but also in key swing states that are likely to be decisive in determining the electoral college outcome.

@@

For the purposes this analysis, a subset of variables was selected from the raw data set. Two additional variables, 'national' and 'state', were created using the existing 'state' and 'end date' variable. A short description of each variable of interest and their respective measurement methods is given below.

@@

@tbl-sumstat shows that the average poll in our dataset has a numeric grade of 2.2. This suggests most polls are of moderate to high quality in terms of reliability. The mean pollscore of -0.5 suggests these polls are relatively accurate, as negative values imply reduced bias. Additionally, the large average sample size of 1908 respondents across polls provides a stable basis for our model, reducing variability and ensuring reliable predictions for candidate support.

@@

@tbl-topster displays the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality). We observe that the most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, indicating it provides fairly reliable data. However, there is variability in pollscore and numeric grade across the top pollsters, reflecting differences in quality and potential biases among them, which the model needs to adjust for to achieve accurate predictions.

@@

@fig-ster shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations. Morning Consult, for example, demonstrates a wide range in reported support, with more spread in Trump’s support. In contrast, Siena/NYT shows less variation, with Harris consistently leading. This variability across pollsters emphasizes the importance of aggregating multiple polls to account for organization-specific biases and ensure a more balanced view of candidate support.

@@

@fig-ss shows a clear right-skewed distribution. Most of the sample sizes are clustered between 0 and 3000 respondents, with a sharp peak around 1000-1500 respondents. This indicates that the majority of polls have smaller sample sizes. As sample size increases, the frequency significantly drops, with very few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more. This wide range in sample sizes can affect the precision of estimates across different polls.

@@

@fig-reli shows the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations. The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, indicating a concentration of pollsters with moderate to high reliability scores. In contrast, the Pollscore distribution, where lower values indicate higher reliability, shows a range primarily between -1.5 and 0, with a notable peak around -0.5. This suggests that while many polls demonstrate relatively low bias, there is still variability in reliability across organizations. The distinction between these two metrics emphasizes the need to consider both quality (Numeric Grade) and potential systematic bias (Pollscore) when weighting polls in the model.

@@

In this paper, we set out to predict the support for both Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election using a “polls-of-polls” approach. By aggregating multiple polls, we sought to mitigate the inherent biases present in individual surveys, creating a more accurate and balanced forecast across candidates. Our analysis was driven by a linear regression model using key predictors, such as pollster, sample size, state, and recency of the polls. After calculating predicted values from the model, we applied a weighting scheme based on each pollster’s numeric grade and pollscore, incorporating both reliability and bias. This weighted approach enabled us to produce an overall prediction of candidate support, reflecting variations across different states and polling organizations.

@@

In conclusion, this paper contributes to election forecasting by providing a comprehensive, weighted estimate of candidate support across both national and state levels. While our predictions provide valuable insights into potential election outcomes, the evolving nature of voter sentiment and poll accuracy reflects the need for adaptive models that account for ongoing changes in public opinion.

>>>>

While individual polls provide snapshots of public opinion, they are often subject to biases and methodological differences. In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by using data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy. Our analysis focuses on estimating the levels of support for each candidate nationally while focusing on key swing states, which are likely to be decisive in determining the electoral college outcome.

@@

For this analysis, a subset of variables was selected from the raw data set. Two additional variables, 'national' and 'state', were created using the existing 'state' and 'end date' variable. A short description of each variable of interest and their respective measurement methods is given below.

@@

@tbl-sumstat shows the average poll in our data set has a numeric grade of 2.2. This suggests most polls are of moderate to high quality in terms of reliability. The mean pollscore of -0.5 suggests these polls are accurate, as negative values imply reduced bias. Also, the large average sample size of 1908 respondents across polls reduces variability and ensures reliable predictions for the candidate support model.

@@

@tbl-topster shows the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality). The most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, suggesting the data provided is reliable. However, there is variability in pollscore and numeric grade across the top pollsters, showing differences in quality and potential biases among them.

@@

@fig-ster shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations. Morning Consult, for example, demonstrates a wide range in reported support, with more spread in Trump’s support. In contrast, Siena/NYT shows less variation, with Harris consistently leading. This variability across pollsters shows the importance of aggregating multiple polls to account for organization-specific biases and ensure a more balanced view of candidate support.

@@

@fig-ss shows a clear right-skewed distribution. Most of the sample sizes are clustered between 0 and 3000 respondents, with a sharp peak around 1000-1500 respondents. This indicates that the majority of polls have smaller sample sizes. As sample size increases, the frequency decreases, with few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more. This wide range in sample sizes can affect the precision of estimates across different polls.

@@

@fig-reli shows the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations. The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, suggesting a concentration of pollsters with moderate to high reliability scores. In contrast, the Pollscore distribution, where lower values indicate higher reliability, shows a range primarily between -1.5 and 0, with a notable peak around -0.5. This suggests that while many polls demonstrate relatively low bias, there is still variability in reliability across organizations. The distinction between these two metrics emphasizes the need to consider both quality (Numeric Grade) and potential systematic bias (Pollscore) when weighting polls in the model.

@@

In this paper, we set out to predict the support for both Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election using a “polls-of-polls” approach ADD REFERENCE HERE. By aggregating multiple polls, we sought to mitigate the inherent biases present in individual surveys, creating a more accurate and balanced forecast across candidates. Our analysis was driven by a linear regression model using key predictors, such as pollster, sample size, state, and recency of the polls. After calculating predicted values from the model, we applied a weighting scheme based on each pollster’s numeric grade and pollscore, incorporating both reliability and bias. This weighted approach enabled us to produce an overall prediction of candidate support, reflecting variations across different states and polling organizations.

@@

In conclusion, this paper contributes to election forecasting by providing a weighted estimate of candidate support across both national and state levels.The evolving nature of voter sentiment and poll accuracy suggests the need for adaptive models that takes in consideration ongoing changes in public opinion.
